comgen-outline

### 2017-12-05 paper/comgen-paper.Rmd
edits to four sections of discussion

## The Comparison Class and the Question Under Discussion

In this paper, we proposed a model for understanding generalizations in language that relies upon interlocutors' shared beliefs over the statistics of the event or property under discussion.
We constructed the prior belief distribution for the probability of a property, event, or cause by considering other possible categories having the property, people doing the action, or causes producing the effect.
These other kinds, people, and causes form *comparison classes* against which the referent-category is evaluated. 
Thus, the feature-probability prior $P(h)$ is always relative to a class of other entities or events, a comparison class $C$: $P_C(h)$.

The existence of comparison classes is uncontroversial in the study of *vague language* [e.g., gradable adjectives like *tall* and vague quantifiers like *many*; @Bale2011; @Solt2009].
Adult judgments of the felicity of gradable adjectives like *tall* or *dark* depend upon fine-grained details of the statistics of the comparison class [@Qing2014; @Schmidt2009; @Solt2012].
We have shown similar sensitivity to the statistics of comparison class in the language of generalizations, thus incorporating generic, habitual, and causal language into the study of *vague language*.

In this work, we made the *a priori* decision to construct the comparison class with respect to other kinds, people, or causes.
In this way, the statement of genericity is addressing the implicit Question Under Discussions (QUDs): *who \textsc{does action}?*, *what causes \textsc{effect}?*, or *what \textsc{has feature}?*, respectively.
In the minimal contexts employed in our experiments, this is a reasonable assumption.
There is a different assumption we could have made, however: Another clear way to construct the comparison class is with respect to other features. 
This *feature-wise* comparison class would correspond to addressing the implicit QUDs: *what does \textsc{person} do?*, *what effects does \textsc{cause} bring about?*, and *what features does \textsc{kind} have?*

Statements of genericity can be used to address either QUD, and this difference can even be observed in the same sentence.
The statement "Lawyers care about the law." can be interpreted with a category-wise comparison class (i.e., *Lawyers (as opposed to doctors, firefighters, ...) care about the law.*; one can imagine this interpretation in a pedagogical context) or a feature-wise comparison class (i.e., *Lawyers care about the law (as opposed to justice, ethics, ...)* e.g., when said sardonically).
We chose to focus in this work on the category-wise comparison class for methodological convenience. 
In our three case studies, the *category-wise* comparison classes are fixed to *other people*, *other causes*, or *other kinds*, even as we vary the features.
*Feature-wise* comparison classes are likely to be modulated by the feature itself.
Future work should address the feature-wise reading of generalizations, and explore what cues lead listeners to one or the other interpretation.

@@

## Structured prior knowledge and implications for conceptual accounts

@@

Previous psychological and philosophical work on generics has looked beyond feature-probability and focused on conceptual distinctions and relations [@Gelman2003; @Prasada2013; @Leslie2007; @Leslie2008].^[
In these accounts, feature-probability is normally treated as synonymous with the current *prevalence* of the feature in the category. 
As we have shown, feature-probability in our account is more general notion, reflecting speaker's subjective, predictive probabilities.
] 
Prasada has argued for a distinction between \emph{characteristic} properties (e.g., \emph{Diapers are absorbent.}) and \emph{statistical} properties (e.g., \emph{Diapers are white.}).
Leslie suggests information that is striking (e.g., \emph{Tigers eat people.}) is useful and thus permitted to be a generic.
Gelman outlines how generics tend to express \emph{essential} qualities that are relatively timeless and enduring. 
Where in the probability-based semantics could such conceptual distinctions come into play?

Our approach makes the strong claim that beliefs about predicted feature-probability is the connective tissue between conceptual knowledge and the semantics of generic language.
That is, the effect of conceptually meaningful differences on generic language is predicted to be mediated by differences in corresponding feature-probability distributions.
We found in Expts. 1 \& 2 that conceptually different kinds of properties and events gave rise to different feature-probability priors. 
In Expt. 3 we showed that these feature-probability distributions are causally related to endorsement.

It is natural to ask how subjective probabilities might reflect conceptual knowledge.
We found that empirical feature-probability distributions are structured in a way that reflects intuitions about causal mechanisms underlying different properties; the differences in shape of these distributions in turn led to variable endorsements of generalizations. 
It is plausible that richer conceptual knowledge also influences these distributions, such as higher-order conceptual knowledge about the nature of properties and categories [@Gelman2003; @Keil1992].
Indeed, it has been argued that conceptual structure in general, including higher-order abstractions, can be captured by probabilisitic causal models [@pearl1988probabilistic; @Gopnik2003theory] and their generalization in probabilistic programs [@Goodmanconcepts].
Future work will be needed to explore whether probabilistic representations of conceptual knowledge can capture the relations identified in other accounts of generics (such as principled, essential, and striking properties), and whether the effect of these relations can then be adequately captured via their impact on subjective feature-probability.

## Communicating Predictive Probabilities

It is important to note that our approach is based on \emph{subjective} probability, and not mere frequency.
Indeed, we elucidated in Expts. 2c \& 2d that using participants' \emph{predictions} of probability in our formal model perfectly track generic endorsement, when the present frequency would make the wrong prediction.

This paper puts forth the theory that the simplest language of abstractions (i.e., the language of generalizations) communicates predictive probabilities from speaker to hearer. 
This might seem contrary to a basic tenet in cognitive psychology, that human cognition falters when reasoning about probabilities [@Tversky1974].
In a multitude of demonstrations, Kahneman and Tversky observed that human reasoners perform poorly when reasoning about explicit probabilities.
This has been taken to imply that people are bad with probabilties.

We suggest that problems with communicating probabilities are a problem of the language used to convey *explicit* probabilties [cf., @Levinson1995]. 
A host of developmental evidence suggests that even the youngest learners are actually quite good about reasoning about probabilities, but in ways that are not explicit or tied to language [@Gweon2010; @Dewar2010].
In this work, we have argued that probabilities are central to even the most basic linguistic expressions.
We convey probabilities to each other using language, but the most basic ways we have of conveying them are *vague*.

@@

The linguistic outlet for generalizations about categories---so called, *generic language*---has received tremendous attention from psychologists, linguists, and philosophers.
Generic language is one of the earliest emerging forms of complex, compositional language.
Somewhere between 2 and 3 years of age, children recognize that generics convey a generalization about a category, not directly tied to concrete instances in a scene [@Cimpian2008].
Generics are ubiqiutous in child-directed and child-produced speech [@Gelman1998; @GelmanEtAl2004] and are believed to be central to the growth of conceptual knowledge [@Gelman2004],

What is perhaps surprising from a formal perspective is that generic language is far from trivial to characterize. 
If generics convey something about the probabilty of a feature, we would expect them to be in some way comparable to quantifier statments (e.g., "some", "most", "all", ...).
The extreme flexibility of generic meaning (e.g., *Birds lay eggs* vs. *Birds are female*) stands in stark contrast to its early emergence in development.
In fact, quantified statements, whose formal meaning is much more straight-forward (e.g., "Some" means more than zero), emerge *later* in development [@Brandone2014; @Gelman2015].
This has led some to conclude that the normal tools for describing the semantics of quantified utterances (i.e., a truth-functional threshold) is not approrpriate for generic language [@Leslie2008].

We think this would be throwing out the baby with the bathwater.
When we consider the acquisition problem, there are two aspects of meaning that a language learner must acquire for the semantics of a quantified utterance (e.g., "some", "most"): (1) that the meaning is a threshold-function (i.e., $\denote{u}(h, \theta) = \{h :h > \theta\}$) and (2) the fixed value of the threshold (e.g., $\theta = 0$ for "some", $\theta = 0.5$ for "most").
Upon learning the meaning function is a threshold-function (1), a Bayesian language learner would be well-suited to adopt a prior distribution over possible thresholds $\theta$.
This prior could then be updated with more data to acquire its fixed meaning.
For the generic statement, the language learner is done once she acquires (1). 
The rest of the meaning is inferred in context.
With that lens, generic meaning would be primary, while quantified utterances would be refinements.

There is an alternative reason why our uncertain threshold model makes the acquisition problem easier than quantifiers.
A uniform prior over the threshold in the truth-functional threshold semantics (as we have assumed in this work) is mathematically equivalent to a *soft semantics* wherein the degree to which the utterance is true is proportional to the probability degree $h$ *itself*: ^[It is equivalent to the probability degree $h$ normalized by the prior distribution over the degree (i.e., the comparison class)]
$$
\int_{0}^{1} \delta_{h > \theta} \diff \theta =  h
$$

This *soft semantics* (corresponding intuitively to something like "the higher the probability, the better the utterance", or just "more is better") is perhaps the simplest semantics one could imagine.
<!-- Given this mathematical equivalence, we can imagine the *more is better* soft semantics is what is acquired early. -->
The difficulty in acquiring the meaning of quantifiers is then a difficulty in recognizing a fixed-threshold semantics as a special case of this *more is better* semantics.
This special case involves learning the value of the threshold (aspect 2) and not in recognizing that the utterance conveys something about the quantity (e.g., aspect 1).

## Relationship to Other Theories

@@

### Statistical Accounts

@@

### Conceptual Accounts

@@

## Other philosophical puzzles

1. Books are paperbacks.
2. Mary handles the mail from Antarctica.
3. Elephants live in Africa and Asia.
4. Women are submissive.

>>>>

## Implicit aspects of language use: The comparison class and question under discussion

In this paper, we proposed a model for understanding generalizations in language that formalizes a simple semantic theory that interfaces with interlocutors' shared beliefs over the statistics of the event or property in question, represented by the prior belief distribution over the feature-probability.
The feature-probability priors for a property, event, or cause are constructed by considering other possible categories having the property, people doing the action, or causes producing the effect.
Collectively, these other kinds, people, or causes form classes against which the referent-category is evaluated.
That is, the feature-probability prior $P(h)$ is constructed with respect a class of other entities or events, *the comparison class* $C$: $P_C(h)$.

The existence of comparison classes is uncontroversial in the study of *vague* or *underspecified* language [e.g., gradable adjectives like *tall* and vague quantifiers like *many*; @Bale2011; @Solt2009].
Adult judgments of the felicity of gradable adjectives like *tall* or *dark* depend upon fine-grained details of the statistics of the comparison class [@Qing2014; @Schmidt2009; @Solt2012] in ways directly analagous to the context-sensitivity of the language of generalizations explored in this paper.
Thus, we conclude generic, habitual, and causal language can be seen as a special case of *vague language*.

The construction of a comparison class adds another layer of flexibility into the use of generalizations in language. 
Objects can be conceptualized and categorized in multiple ways, giving rise to multiple, logically-possible comparison classes (e.g., a robin is a bird, is an animal, is an object). 
We have begun doing work investigating how pragmatic reasoning can help constrain this open-ended problem [@Tessler2017].
Even if interlocutors coordinate on a particular comparison class, they may disagree about the feature-probabilities of other *other categories* in the class.
That is, our model makes the prediction that the acceptance or rejection of the statement "Humans cause global warming" or "Muslims are terrorists" depends upon a speaker's beliefs about how *other forces* (e.g., plate tectonics) influence the climate or how often other social categories (e.g., neo-nazis) commit acts of terrorism.
Indeed, we saw in Expt. 3 how the statistics of other causes can influence the endorsement of a single generalization (e.g., "Herb C makes animals sleepy").
Understanding how knowledge of the comparison class influences language use is an exciting new area for future research in the language of generalizations.

Rational models of communication, of the kind we have developed in this paper, are required to make explicit a speaker's assumptions about the goal of communication, otherwise known as the Question Under Discussion or QUD [@roberts1996qud]. 
We have made a very minimalist assumption, that the goal of communication (at least, for purposes of endorsement) is the feature-probability.
However, the choice in constructing a comparison class can be thought of as the result of addressing a particular QUD.
For example, the comparison classes used in our case studies were with respect to the target category (*other animals*, *other people*, *other possible causes*); thus, our model addresses the implicit QUDs of "*what* has feature?", "*who* does action?", or "*what* causes effect?".
Were the QUDs to be reversed (e.g., "what features does category have?", "what does person do?", "what effects does cause bring about?"), the comparison classes would need to be constructed with respect to the feature (e.g., *other features*, *other actions*, *other effects*).

The QUD is thus another layer of flexibility that makes understanding generalizations so challenging.
The same sentence can be used to address multiple QUDs [@KrifkaGenericBookFocus].
The statement "Lawyers care about the law." can be interpreted with a category-wise comparison class (i.e., "Lawyers *(as opposed to doctors, firefighters, ...)* care about the law."; imagine in a pedagogical context) or a feature-wise comparison class (i.e., "Lawyers care about the law *(as opposed to justice, doing what's morally right, ...)*" e.g., when said sardonically).
In this work, we focused on category-wise comparison classes for methodological convenience. 
<!-- In our three case studies, the *category-wise* comparison classes are fixed to *other people*, *other causes*, or *other kinds*, even as we vary the features. -->
<!-- *Feature-wise* comparison classes are likely to be modulated by the feature itself. -->
Future work should investigate the feature-wise reading of generalizations, and explore what cues listeners adopt for one or the other interpretation.

<!-- In the minimal contexts employed in our experiments, this is a reasonable assumption. -->
<!-- There is a different assumption we could have made, however: Another clear way to construct the comparison class is with respect to other features.  -->
<!-- This *feature-wise* comparison class would correspond to addressing the implicit QUDs: *what does \textsc{person} do?*, *what effects does \textsc{cause} bring about?*, and *what features does \textsc{kind} have?* -->

@@

## Structured conceptual prior knowledge

@@

Previous psychological and philosophical work on generics has looked beyond feature-probabilities and focused on conceptual distinctions and relations [@Gelman2003; @Prasada2013; @Leslie2007; @Leslie2008].
Prasada has argued for a distinction between \emph{characteristic} properties (e.g., "Diapers are absorbent") and \emph{statistical} properties (e.g., "Diapers are white").
Leslie suggests information that is striking (e.g., "Tigers eat people") is ecologically useful and thus permitted to be a generic.
Gelman outlines how generics tend to express \emph{essential} qualities that are relatively timeless and enduring. 
Where in the probability-based semantics could such conceptual distinctions come into play?

Our approach makes the strong claim that beliefs about predicted feature-probability are the connective tissue between conceptual knowledge and the semantics of generic language.
That is, the effect of conceptually meaningful differences on generic language is predicted to be mediated by differences in corresponding feature-probability distributions.
Modulations of feature-probability distributions can occur for multiple reasons (e.g., by changing the comparison class) as described in the previous section.
We found in Expts. 1 \& 2 that conceptually different kinds of properties and events gave rise to quantitatively different feature-probability priors, even while keeping the comparison class fixed. 
In Expt. 3 we showed that these feature-probability distributions are causally related to endorsement.

It is natural to ask how feature-probabilities might reflect conceptual knowledge.
We found that empirical feature-probability distributions are structured in a mixture-model that plausibly reflects intuitions about causal mechanisms underlying different properties; the differences in shape of these distributions in turn led to variable endorsements of generalizations. 
It is plausible that richer conceptual knowledge also influences these distributions, such as higher-order conceptual knowledge about the nature of properties and categories [@Gelman2003; @Keil1992].
Indeed, conceptual structure in general, including higher-order abstractions, can be captured by probabilisitic causal models [@pearl1988probabilistic; @Gopnik2003theory] and their generalization in probabilistic programs [@Goodmanconcepts].
Future work will be needed to explore whether probabilistic representations of conceptual knowledge can capture the relations identified in other accounts of generics (such as principled, essential, and striking properties), and whether the effect of these relations can then be adequately described via their impact on feature-probability priors.

## Communicating predictive probabilities

This paper puts forth the theory that the language of generalization communicates predictive probabilities from speaker to hearer.
In previous accounts, feature-probability is treated as synonymous with the prevalence or frequency of the feature in the category. 
We make the important distinction that probability can be more general than frequency, and in our account, underlies speakers' subjective, predictions about the future and is subject to top-down moderator.
Expts. 2c \& 2d that showed using participants' \emph{predictions} about the future as the object of communication in our formal model perfectly tracked habitual endorsement, when the present frequency would make the wrong prediction.

Communicating probabilities might seem contrary to a basic tenet in cognitive psychology, that human cognition falls short when reasoning about probabilities [@Tversky1974].
Our theory suggests that the problems with understanding probabilities observed in classic, cognitive psychology paradigms are a problem in understanding the *language used* to convey *explicit probabilties* [cf., @Levinson1995]. 
Rather than conveying *explicit probabilities*, generalizations we argue convey *implicit probabilities*, which are easier to understand and quite useful for statistical reasoning.
Given the relationship to *vague language* (described above), we argue that the utterance "70% of birds fly" is analagous to saying that "John is 6'3"", while saying "Birds fly" is more similar to saying "John is tall".
The latter statements are easier to process, understood at an earlier age, and may be more useful for human reasoning. 
Indeed, young children are actually quite good about reasoning about probabilities, but in ways that are not explicit or tied to language [@Gweon2010; @Dewar2010].

@@

The linguistic outlet for generalizations about categories---so called, *generic language*---has received tremendous attention from psychologists, linguists, and philosophers.
Generic language is one of the earliest emerging forms of complex, compositional language.
Somewhere between two and three years of age, children recognize that generics convey a generalization about a category, not directly tied to concrete instances in a scene [@Cimpian2008].
Generics are common in child-directed and child-produced speech [@Gelman1998; @GelmanEtAl2004] and are believed to be central to the growth of conceptual knowledge [@Gelman2004],

What is perhaps surprising from a formal perspective is that generic language is far from trivial to characterize. 
If generics convey something about the probabilty of a feature, we would expect them to be in some way comparable to quantifier statments (e.g., "some", "most", "all", ...).
The extreme flexibility of generic meaning (e.g., "Birds lay eggs" vs. "Birds are female") stands in stark contrast to its early emergence in development.
In fact, quantified statements, whose formal meaning is much more straight-forward (e.g., "Some" means more than zero), emerge *later* in development [@Brandone2014; @Gelman2015], and are often confused for generic statements [@Leslie2012a].
This has led some to conclude that the normal tools for describing the semantics of quantified utterances (i.e., a truth-functional threshold) is not approrpriate for generic language [@Leslie2008].

We think rejecting the normal tools of truth-functional semantics for generics would be throwing out the baby with the bathwater.
When we consider the acquisition problem, there are three aspects of meaning that a child must acquire for a truth-functional quantifier semantics $\denote{u}(h, \theta) = \{h :h > \theta\}$: 
(1) the dimension being described (i.e., the probability $h$),
(2) the polarity of the relation (i.e., $>$ vs. $<$), and 
(3) the value of the threshold (e.g., $\theta = 0$ for "some", $\theta = 0.5$ for "most").
Upon learning the meaning has to do with probability and that it stands in a positive relation to some standard $\theta$, a rational learner would adopt a prior distribution over possible thresholds $\theta$, as a form of *lexical uncertainty* [@Bergen2016].
This prior could then be updated with more data to acquire a fixed meaning.
However, for the generic statement, the language learner is done once she acquires aspects (1) and (2).
In our model, $\theta$ is left underspecified in the semantics and is inferred in context.
Thus, in this sketch of an acquisition model, a child would first acquire generic language, and from that, carve out more precise meanings for quantified language.

The uniform prior over the threshold $\theta$ in the truth-functional threshold semantics (as we have assumed in our model) provides a secondary argument for why acquiring our semantics for generics should be easy.
Uniform uncertainty over values in the unit interval $[0, 1]$ for threshold $\theta$ in a threshold-function is mathematically equivalent to a *soft semantics* wherein the degree to which the utterance is true is proportional to the feature-probability $h$ *itself*: 
$$
\int_{0}^{1} \delta_{h > \theta} \diff \theta =  h
$$

This *soft semantics* (corresponding intuitively to a meaning like "the higher the probability, the better the utterance", or simply "more is better") is perhaps the simplest semantics one could imagine.
<!-- Given this mathematical equivalence, we can imagine the *more is better* soft semantics is what is acquired early. -->
The difficulty in acquiring the meaning of quantifiers is then a difficulty in recognizing a fixed-threshold semantics as a special case of this *more is better* semantics.
We leave for future work the precise implementation of the acquisition model we sketch out here.
<!-- This special case involves learning the value of the threshold (aspect 2) and not in recognizing that the utterance conveys something about the quantity (e.g., aspect 1). -->
## Other philosophical puzzles

1. Books are paperbacks.
2. Mary handles the mail from Antarctica.
3. Supreme Court Justices have even social security numbers.
4. Elephants live in Africa and Asia.
5. Women are submissive.

## Relationship to other theories

@@

### Statistical accounts

@@

### Conceptual accounts

@@
